### 2016-04-20 ALMA/Cycle4/RXJ.tex
minor rewording

\begin{center}
{\large{\bf
{ 
Molecular gas dynamics and AGN/starburst mechanism in a
strongly-lensed wet merger: bridging the gap between local ULIRGs and high-z systems}
}}
\end{center}
\vspace{-0.545em}
\centerline{\normalsize{\bf PI: 
{T. K. Daisy Leung}}} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{0.1em} {\bf Missing link between mergers/ULIRGs and their high-z analogues} 
Ultraluminous infrared galaxies (ULIRGs: \LIR$\geq$10$^{12}$\Lsun) have been
regarded as analogues of high-redshift (z) starbursts given the
similarities in their interstellar medium (ISM; e.g. \LIR/\Lp).
As such, detailed studies 
of ULIRGs are important 
to gain more detailed insight into the early universe and in studies of galaxy evolution over cosmic time.
Since mergers are believed to play an important role in giving rise to these dusty galaxies
(e.g. Sanders \& Mirabel 1996). Yet, merger-induced effects on the physical mechanisms and chemistry 
that drive the 
%intense 
starburst (SB) and active galactic nucleus (AGN) activities on small scales are still unclear.
Thus characterizing the properties of the molecular gas that fuels star-formation (SF) and AGN is
crucial to understand the interplay between AGN and SB
and their relation to the ISM properties of galaxies across cosmic time.

@@

% 
% mounting supporting evidence for such elevated
% HCN emission at the centers of local Seyfert galaxies has been accumulated (e.g., Kohno
% et al. 2001; Imanishi et al. 2006; Krips et al. 2008), giving an empirical AGN diagnostic.
%
%mounting evidence that HCN lines can be
%overluminous with respect to the lines of other dense molecular
%gas tracers, like HCO+, in the circumnuclear disks of
%Seyferts (Kohno et al. 2001; Usero et al. 2004; Kohno 2005).
%A significant percentage of luminous and ultraluminous infrared
%galaxies (LIRGs and ULIRGs) has also been reported
%to show high HCN/HCO+ intensity ratios (Graciá-Carpio et al.
%2006; Imanishi et al. 2006, 2007).
% 
%  More recently, spatially resolved ALMA observations
% of HCN(4-3) and HCO+(4-3) at the CND of NGC 1068 at a ?50 pc resolution reveal
% that the HCN(4-3)/HCO+(4-3) line ratios are indeed highly elevated at the CND but
% not significantly at the very position of the active nucleus (Garc??a-Burillo et al. 2014;
% Viti et al. 2014). This indicates that HCN enhancement is not simply caused by X-ray
% irradiation. Recent HCN(4-3)/HCO+(4-3) measurements of the high-luminosity
% type-1 AGN NGC 7469 (Izumi et al. 2015a) with LX = 2 × 1043 erg s?1
% suggest that the
% degree of HCN enhancement is not very different from that in NGC 1097 (Izumi et al.
% 2013), a much lower luminosity type-1 AGN (LX = 7×1040 erg s?1
% ), also adding further
% evidence against the XDR origin of enhanced HCN in AGNs (see also Mart??n et al. 2015).
% 
% XDR versus PDRs
Since \eco is excited in higher temperature than the HCN and \hcop lines (e.g. in 
X-ray dominated regions (XDRs) near an AGN), 
variations in HCN/CO will allow us to constrain the spatial extent of photon-dominated regions
 (PDRs from SBs) and XDRs.
% physical properties
In addition, we will measure spatially resolved line ratios of HCN/CO and \hcop/CO within the AGN host galaxy
as proxies of its very dense (cores; $\sim$10$^5$-10$^7$ cm$^{-3}$) versus the less dense (clumps; $\sim$10$^4$
cm$^{-3}$) gas content and their spatial distributions as a function of distance from the AGN/nucleus. This will enable
us to gain insight into the interplay between AGN, SB and the dense gas content in the ISM (and thus the SF mode on small scales across different galaxy populations as traced by the HCN/CO-\LIR relation).
% \Fig{SF}).
% CO SLED
We will also combine \eco with our existing \bco and CO($J$=3\rarr2) data to constrain the gas density (n$_{H_{2}}$) 
and kinetic temperature by performing large velocity gradient modeling.

\underline{\bf The SK Law for dense gas:}
The Schmidt-Kennicutt (SK) relation (%``Star-formation law"; 
$\Sigma_{\rm SFR}$$\propto$$\Sigma^N_{\rm gas}$)
is one of the key ingredients
for theoretical models as it encodes the physical processes and timescales
regulating star-formation and their dependence on the ISM 
(e.g. \citealt{Narayanan14a}).
However, the surface density of gas at high densities (\ncrit$\gtrsim$10$^5$cm$^{-3}$; 
$\Sigma_{\rm dense}$)
% ; and thus $\Sigma_{\rm SFR}\propto \Sigma^N_{\rm dense}$) 
is currently poorly constrained 
for high-z galaxies due to the lack of 
(spatially resolved) observations of the much weaker emission 
from high critical density tracers. %(see right of \Fig{SF}).
It is thus unclear how the power-law index of the SK law should change depending on
the critical density of the tracer used to probe the SF gas \citep[e.g.][]{KT07a}
and how the SFR surface density should depend on the global 
dynamical time scales in normal galaxies and in mergers.
At the proposed resolution, we will spatially resolve, for the first time,
the {\it true SK relation} by measuring the dense gas surface density (\ncrit$\gtrsim$10$^5$cm$^{-3}$)
in a high-z merger, down to the size scale of star-forming gas clumps. 
Hence, providing crucial constraints on star-formation at high redshift.

\noindent Our proposed observations therefore provide an exceptional opportunity to investigate
the physical properties and dynamical structures of different gas phases in the ISM of mergers 
and distant galaxies with exquisite details. Such investigations are vital to bridge the gap in our 
current understanding on galaxy populations across cosmic time.
This will also demonstrate the capabilities of ALMA in % mapping 
utilizing these much fainter emission 
high-dipole moment molecules as routine tracers to 
study star-formation at earlier epochs.

\begin{figure*}[!tbph]
\vspace{-1.85em}
\centering
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=0.5575\textwidth}}]{figure}[\FBwidth]
{
% HCN/CO - LIR plot: e.g. Fig 6 in \citep{Juneau09a}
\includegraphics[trim=15 0 5 10, clip, width=0.257\textwidth]{Fig/SFRSD}} 
{
\hspace{-0.54em}
\caption{ \fontsize{10pt}{12pt}\selectfont{
% \textbf{HCN/CO-\LIR relation} % 
%An increase in line ratios between high-density and low-density molecular gas tracers (e.g. HCN/CO) 
%with increasing infrared luminosity suggests that the main driver of the enhanced SFR
%is the molecular gas density distribution, in agreement with the observations that ULIRGs
%host large reservoirs of dense molecular gas in their
%central regions. 
\textbf{Schmidt-Kennicutt relation for dense gas}
While constraints on the power-law index from gas of different densities are important for 
star-formation models, current studies only have constraints
up to the gas of density traced by the ground state HCN line (\ncrit$\sim$10$^{4}$),
which are also largely limited to local measurements \citep{GB12a}. 
Our proposed observations will, for the first time, spatially resolve the 
SK relation at kpc scale 
in a ULIRG at $z$$\sim$0.7, using
high critical density tracers (\dhcn and \dhcop). This will allow us to 
explore potential deviations at higher redshift and provide
key constraints for models of galaxy evolution.
% This will also demonstrate the utility of mid-$J$ lines in tracing 
% the star-forming dense gas, which will be routinely mapped in high-z galaxies with ALMA.
}
\label{fig:SF}}
}
\vspace{-0.905em}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent {\bf Technical overview}
We estimate the source size from our lens model 
and adopt conservative line ratios measured for ULIRGs and high-z galaxies 
(\citealt[][]{Greve09a,CW13}; GS04; P07) to compute the estimate line fluxes.
To secure enough S/N for lens modeling, 
we require a minimum of 8$\sigma$ of 0.47 mJy beam\pmOne and 0.07 mJy beam\pmOne
per 150 \kms channel 
for the science goals, respectively (see TJ for details).

>>>>

\begin{center}
{\large{\bf
{ 
Molecular gas dynamics and AGN/starburst mechanism in a
strongly-lensed wet merger: bridging the gap between local ULIRGs and high-z systems}
}}
\end{center}
\vspace{-0.545em}
\centerline{\normalsize{\bf PI: 
{T. K. Daisy Leung}}} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{0.1em} {\bf Missing link between mergers/ULIRGs and their high-z analogues} 
Ultraluminous infrared galaxies (ULIRGs: \LIR$\geq$10$^{12}$\Lsun) have been
regarded as analogues of high-redshift (z) starbursts given the
similarities in their interstellar medium (ISM; e.g. \LIR/\Lp).
As such, detailed studies 
of ULIRGs are important 
to gain more detailed insight into the early universe and in studies of galaxy evolution over cosmic time.
While mergers are believed to play an important role in giving rise to these dusty galaxies
(e.g. Sanders \& Mirabel 1996), merger-induced effects on the physical mechanisms and chemistry 
that drive the 
%intense 
starburst (SB) and active galactic nucleus (AGN) activities on small scales are still unclear.
Thus characterizing the properties of the molecular gas that fuels star-formation (SF) and AGN is
crucial to understand the interplay between AGN and SB
and their relation to the ISM properties of galaxies across cosmic time.

@@

% 
% mounting supporting evidence for such elevated
% HCN emission at the centers of local Seyfert galaxies has been accumulated (e.g., Kohno
% et al. 2001; Imanishi et al. 2006; Krips et al. 2008), giving an empirical AGN diagnostic.
%
%mounting evidence that HCN lines can be
%overluminous with respect to the lines of other dense molecular
%gas tracers, like HCO+, in the circumnuclear disks of
%Seyferts (Kohno et al. 2001; Usero et al. 2004; Kohno 2005).
%A significant percentage of luminous and ultraluminous infrared
%galaxies (LIRGs and ULIRGs) has also been reported
%to show high HCN/HCO+ intensity ratios (Graciá-Carpio et al.
%2006; Imanishi et al. 2006, 2007).
% 
%  More recently, spatially resolved ALMA observations
% of HCN(4-3) and HCO+(4-3) at the CND of NGC 1068 at a ?50 pc resolution reveal
% that the HCN(4-3)/HCO+(4-3) line ratios are indeed highly elevated at the CND but
% not significantly at the very position of the active nucleus (Garc??a-Burillo et al. 2014;
% Viti et al. 2014). This indicates that HCN enhancement is not simply caused by X-ray
% irradiation. Recent HCN(4-3)/HCO+(4-3) measurements of the high-luminosity
% type-1 AGN NGC 7469 (Izumi et al. 2015a) with LX = 2 × 1043 erg s?1
% suggest that the
% degree of HCN enhancement is not very different from that in NGC 1097 (Izumi et al.
% 2013), a much lower luminosity type-1 AGN (LX = 7×1040 erg s?1
% ), also adding further
% evidence against the XDR origin of enhanced HCN in AGNs (see also Mart??n et al. 2015).
% 
% XDR versus PDRs
Since \eco is excited in higher temperature than the HCN and \hcop lines (e.g. in 
X-ray dominated regions (XDRs) near an AGN), 
variations in HCN/CO will allow us to constrain the spatial extent of photon-dominated regions
 (PDRs from SBs) and XDRs.
% physical properties
In addition, we will measure spatially resolved line ratios of HCN/CO and \hcop/CO within the AGN host galaxy
as proxies of its very dense (cores; $\sim$10$^5$-10$^7$ cm$^{-3}$) versus the less dense (clumps; $\sim$10$^4$
cm$^{-3}$) gas content and their spatial distributions as a function of distance from the AGN/nucleus. This will enable
us to gain insight into the interplay between AGN, SB and the dense gas content in the ISM (and thus the SF mode on small scale across different galaxy populations as traced by the HCN/CO-\LIR relation).
% \Fig{SF}).
% CO SLED
We will also combine \eco with our existing \bco and CO($J$=3\rarr2) data to constrain the gas density (n$_{H_{2}}$) 
and kinetic temperature by performing large velocity gradient modeling.

\underline{\bf The SK Law for dense gas:}
The Schmidt-Kennicutt (SK) relation (%``Star-formation law"; 
$\Sigma_{\rm SFR}$$\propto$$\Sigma^N_{\rm gas}$; \Fig{SF})
is one of the key ingredients
for theoretical models as it encodes the physical processes and timescales
regulating star-formation and their dependence on the ISM 
(e.g. \citealt{Narayanan14a}).
However, the surface density of gas at high densities (\ncrit$\gtrsim$10$^5$cm$^{-3}$; 
$\Sigma_{\rm dense}$)
% ; and thus $\Sigma_{\rm SFR}\propto \Sigma^N_{\rm dense}$) 
is currently poorly constrained 
for high-z galaxies due to the lack of 
(spatially resolved) observations of the much weaker emission 
from high critical density tracers. %(see right of \Fig{SF}).
It is thus unclear how the power-law index of the SK law should change depending on
the critical density of the tracer used to probe the SF gas \citep[e.g.][]{KT07a}
and how the SFR surface density should depend on the global 
dynamical time scales in normal galaxies and in mergers.
At the proposed resolution, we will spatially resolve, for the first time,
the {\it true SK relation} by measuring the dense gas surface density (\ncrit$\gtrsim$10$^5$cm$^{-3}$)
in a high-z merger, down to the size scale of star-forming gas clumps. 
This will enable us to provide crucial constraints on star-formation at high redshift.

\noindent Our proposed observations therefore provide an exceptional opportunity to investigate
the physical properties and dynamical structures of different gas phases in the ISM of mergers 
and distant galaxies with exquisite details. Such investigations are vital to bridge the gap in our 
current understanding on different galaxy populations across cosmic time.
This will also demonstrate the capabilities of ALMA in % mapping 
utilizing these much fainter
high-dipole moment molecules as routine tracers to 
study star-formation at earlier epochs.

\begin{figure*}[!tbph]
\vspace{-1.85em}
\centering
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=0.5575\textwidth}}]{figure}[\FBwidth]
{
% HCN/CO - LIR plot: e.g. Fig 6 in \citep{Juneau09a}
\includegraphics[trim=15 0 5 10, clip, width=0.257\textwidth]{Fig/SFRSD}} 
{
\hspace{-0.54em}
\caption{ \fontsize{10pt}{12pt}\selectfont{
% \textbf{HCN/CO-\LIR relation} % 
%An increase in line ratios between high-density and low-density molecular gas tracers (e.g. HCN/CO) 
%with increasing infrared luminosity suggests that the main driver of the enhanced SFR
%is the molecular gas density distribution, in agreement with the observations that ULIRGs
%host large reservoirs of dense molecular gas in their
%central regions. 
\textbf{Schmidt-Kennicutt relation for dense gas}
While constraints on the power-law index from gas of different densities are important for 
star-formation models, current studies only have constraints
up to the gas of density traced by the ground state HCN line (\ncrit$\sim$10$^{4}$),
which are also largely limited to local measurements \citep{GB12a}. 
Our proposed observations will, for the first time, spatially resolve the 
SK relation at kpc scale 
in a ULIRG at $z$$\sim$0.7 by using
high critical density tracers (\dhcn and \dhcop). This will allow us to 
explore potential deviations at higher redshift and provide
key constraints for models of galaxy evolution.
% This will also demonstrate the utility of mid-$J$ lines in tracing 
% the star-forming dense gas, which will be routinely mapped in high-z galaxies with ALMA.
}
\label{fig:SF}}
}
\vspace{-0.905em}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent {\bf Technical overview}
We estimate the source size from our lens model 
and adopt conservative line ratios measured for ULIRGs and high-z galaxies 
(GS04; P07; \citealt[][]{Greve09a,CW13}) to compute the estimate line fluxes.
To secure enough S/N for lens modeling, 
we require a minimum of 8$\sigma$ of 0.47 mJy beam\pmOne and 0.07 mJy beam\pmOne
per 150 \kms channel 
for the science goals, respectively (see TJ for details).
